Plant Trait Imputation

### 2024-06-03 code/L1/L1_P5_TropicalAndes_plant_imputetraits.Rmd
update to add additional traits

# Convert the wide dataframe to long format
long_plant_traits <- plant_traits %>%
  gather(TraitName, TraitValue, -species)

# Filter rows with NA values
na_traits <- long_plant_traits %>%
  filter(is.na(TraitValue))

# get a list of species
na_species_list <- unique(na_traits$species)
```

@@

# Separate dataframes from list
Dispersal.Syndrome <- trait_dfs$Dispersal.Syndrome
Fruit.Length <- trait_dfs$Fruit.Length
Plant.Height <- trait_dfs$Plant.Height
Fruit.Type <- trait_dfs$Fruit.Type
Growth.Form <- trait_dfs$Growth.Form
```

```{r}
# Create a list of unique genus and families for each trait
Dispersal.Syndrome_families <- unique(Dispersal.Syndrome$family)
Dispersal.Syndrome_genus <- unique(Dispersal.Syndrome$genus)

Fruit.Length_families <- unique(Fruit.Length$family)
Fruit.Length_genus <- unique(Fruit.Length$genus)

Plant.Height_families <- unique(Plant.Height$family)
Plant.Height_genus <- unique(Plant.Height$genus)

Fruit.Type_families <- unique(Fruit.Type$family)
Fruit.Type_genus <- unique(Fruit.Type$genus)

Growth.Form_families <- unique(Growth.Form$family)
Growth.Form_genus <- unique(Growth.Form$genus)
```

## Retrive trait records for species using BIEN & GIFT
```{r}
# Dispersal Syndrome
Dispersal.Syndrome_families_BIEN <- BIEN_trait_traitbyfamily(family = Dispersal.Syndrome_families, trait = "whole plant dispersal syndrome")
Dispersal.Syndrome_genus_BIEN <- BIEN_trait_traitbygenus(genus = Dispersal.Syndrome_genus, trait = "whole plant dispersal syndrome")

Dispersal.Syndrome_families_GIFT <- GIFT_traits_tax(trait_IDs = "3.3.1")
Dispersal.Syndrome_families_GIFT <- Dispersal.Syndrome_families_GIFT[Dispersal.Syndrome_families_GIFT$family %in% Dispersal.Syndrome_families, ]
```

```{r}
# Fruit Length
Fruit.Length_families_BIEN <- BIEN_trait_traitbyfamily(family = Fruit.Length_families, trait = c("maximum fruit length", "minimum fruit length"))
Fruit.Length_genus_BIEN <- BIEN_trait_traitbygenus(genus = Fruit.Length_genus, trait = c("maximum fruit length", "minimum fruit length"))

@@

@@

add column of trait level (specices, genus, family)
```{r}
# Add new column trait_level to long_plant_trait dataframe
long_plant_traits <- long_plant_traits %>%
  mutate(TraitLevel = ifelse(!is.na(TraitValue), "species", NA))
```

>>>>

# get a list of species
na_species_list <- unique(plant_traits$species)
```

```{r}
length(na_species_list)
```

@@

# Separate dataframes from list
DispersalSyndrome <- trait_dfs$DispersalSyndrome
PlantLifespan_years <- trait_dfs$PlantLifespan_years
FruitMass_mg <- trait_dfs$FruitMass_mg
PlantHeight_m <- trait_dfs$PlantHeight_m
GrowthForm <- trait_dfs$GrowthForm
FruitLength_mm <- trait_dfs$FruitLength_mm
SeedWidth_mm <- trait_dfs$SeedWidth_mm
SeedLength_mm <- trait_dfs$SeedLength_mm
SeedMass_g <- trait_dfs$SeedMass_g
```

## Create a list of unique genus and families for each trait
```{r}
DispersalSyndrome_families <- unique(DispersalSyndrome$family)
DispersalSyndrome_genus <- unique(DispersalSyndrome$genus)
length(DispersalSyndrome_families)
length(DispersalSyndrome_genus)
```

```{r}
PlantLifespan_years_families <- unique(PlantLifespan_years$family)
PlantLifespan_years_genus <- unique(PlantLifespan_years$genus)
length(PlantLifespan_years_families)
length(PlantLifespan_years_genus)
```

```{r}

FruitMass_mg_families <- unique(FruitMass_mg$family)
FruitMass_mg_genus <- unique(FruitMass_mg$genus)

PlantHeight_m_families <- unique(PlantHeight_m$family)
PlantHeight_m_genus <- unique(PlantHeight_m$genus)

GrowthForm_families <- unique(GrowthForm$family)
GrowthForm_genus <- unique(GrowthForm$genus)

FruitLength_mm_families <- unique(FruitLength_mm$family)
FruitLength_mm_genus <- unique(FruitLength_mm$genus)

SeedWidth_mm_families <- unique(SeedWidth_mm$family)
SeedWidth_mm_genus <- unique(SeedWidth_mm$genus)

SeedLength_mm_families <- unique(SeedLength_mm$family)
SeedLength_mm_genus <- unique(SeedLength_mm$genus)

SeedMass_g_families <- unique(SeedMass_g$family)
SeedMass_g_genus <- unique(SeedMass_g$genus)
```

## Retrive trait records for species using BIEN & GIFT
### Dispersal Syndrome
```{r}
DispersalSyndrome_families_BIEN <- BIEN_trait_traitbyfamily(family = DispersalSyndrome_families, trait = "whole plant dispersal syndrome")
DispersalSyndrome_genus_BIEN <- BIEN_trait_traitbygenus(genus = DispersalSyndrome_genus, trait = "whole plant dispersal syndrome")

DispersalSyndrome_families_GIFT <- GIFT_traits_tax(trait_IDs = "3.3.1")
DispersalSyndrome_families_GIFT <- DispersalSyndrome_families_GIFT[DispersalSyndrome_families_GIFT$family %in% DispersalSyndrome_families, ]
```

```{r}
nrow(DispersalSyndrome_families_BIEN)
length(unique(DispersalSyndrome_families_BIEN$scrubbed_family))
unique(DispersalSyndrome_families_BIEN$trait_value)
```
```{r}
nrow(DispersalSyndrome_genus_BIEN)
length(unique(DispersalSyndrome_genus_BIEN$scrubbed_genus))
unique(DispersalSyndrome_genus_BIEN$trait_value)
```

```{r}
nrow(DispersalSyndrome_families_GIFT)
```

### Fruit Length
```{r}
Fruit.Length_families_BIEN <- BIEN_trait_traitbyfamily(family = Fruit.Length_families, trait = c("maximum fruit length", "minimum fruit length"))
Fruit.Length_genus_BIEN <- BIEN_trait_traitbygenus(genus = Fruit.Length_genus, trait = c("maximum fruit length", "minimum fruit length"))

@@

```{r}
nrow(DispersalSyndrome_families_BIEN)
length(unique(DispersalSyndrome_families_BIEN$scrubbed_family))
unique(DispersalSyndrome_families_BIEN$trait_value)
```

@@

add column of trait level (species, genus, family)
```{r}
# Add new column trait_level to long_plant_trait dataframe
long_plant_traits <- long_plant_traits %>%
  mutate(TraitLevel = ifelse(!is.na(TraitValue), "species", NA))
```
